---
title: "DATA 607 Assignment Week 1"
author: "Ajay Arora"
date: "August 27, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Reference: http://www.programmingr.com/examples/read-csv-web/

## R Markdown
## Read the source file & show some data.
```{r}
mush_subset <- read.csv(file="https://raw.githubusercontent.com/AjayArora35/DATA-607-Assignment-Week-1/master/agaricus-lepiota.data", header=FALSE, sep=",")
head(mush_subset)
```

##Get the data dictionary and load 
V1. consumable: e=edible, p=poisonous

V2. cap-shape: bell=b,conical=c,convex=x,flat=f, knobbed=k,sunken=s

V3. cap-surface: fibrous=f,grooves=g,scaly=y,smooth=s

V4. cap-color: brown=n,buff=b,cinnamon=c,gray=g,green=r, pink=p,purple=u,red=e,white=w,yellow=y

V5. bruises?: bruises=t,no=f

V6. Odor: almond=a,anise=l,creosote=c,fishy=y,foul=f, musty=m,none=n,pungent=p,spicy=s

V7. gill-attachment: attached=a,descending=d,free=f,notched=n

V8. gill-spacing: close=c,crowded=w,distant=d

V9. gill-size: broad=b,narrow=n

V10. gill-color: black=k,brown=n,buff=b,chocolate=h,gray=g, green=r,orange=o,pink=p,purple=u,red=e, white=w,yellow=y

V11. stalk-shape: enlarging=e,tapering=t

V12. stalk-root: bulbous=b,club=c,cup=u,equal=e, rhizomorphs=z,rooted=r,missing=?

V13. stalk-surface-above-ring: fibrous=f,scaly=y,silky=k,smooth=s

V14. stalk-surface-below-ring: fibrous=f,scaly=y,silky=k,smooth=s

V15. stalk-color-above-ring: brown=n,buff=b,cinnamon=c,gray=g,orange=o, pink=p,red=e,white=w,yellow=y

V16. stalk-color-below-ring: brown=n,buff=b,cinnamon=c,gray=g,orange=o, pink=p,red=e,white=w,yellow=y

V17. veil-type: partial=p,universal=u

V18. veil-color: brown=n,orange=o,white=w,yellow=y

V19. ring-number: none=n,one=o,two=t

V20. ring-type: cobwebby=c,evanescent=e,flaring=f,large=l, none=n,pendant=p,sheathing=s,zone=z

V21. spore-print-color: black=k,brown=n,buff=b,chocolate=h,green=r, orange=o,purple=u,white=w,yellow=y

V22. population: abundant=a,clustered=c,numerous=n, scattered=s,several=v,solitary=y

V23. Habitat: grasses=g,leaves=l,meadows=m,paths=p, urban=u,waste=w,woods=d


#Create a subset of the data.  We will use V1 = Consumable, V23 = Habitat, V2 = cap-shape, V6 = Odor, V4 = cap-color, and V10 - gill-color


#Transform headers names using the dictionary.
Reference: http://www.cookbook-r.com/Manipulating_data/Renaming_columns_in_a_data_frame/

```{r}
mush_subset <- data.frame(mush_subset$V1, mush_subset$V2, mush_subset$V4, mush_subset$V6, mush_subset$V10, mush_subset$V23)
head(mush_subset)
```

```{r}
colnames(mush_subset) <- c("Consumable","Cap_Shape", "Cap_Color", "Odor", "Gill_Color", "Habitat")
head(mush_subset)

```
#Substitute data with corresponding data dictionary names
Reference: https://stackoverflow.com/questions/5824173/replace-a-value-in-a-data-frame-based-on-a-conditional-if-statement

```{r}
mush_subset$Consumable <- as.character(mush_subset$Consumable)
mush_subset$Consumable[mush_subset$Consumable == "e"] <- "edible"
mush_subset$Consumable[mush_subset$Consumable == "p"] <- "poisonous"

mush_subset$Cap_Shape <- as.character(mush_subset$Cap_Shape)
mush_subset$Cap_Shape[mush_subset$Cap_Shape == "b"] = "bell"
mush_subset$Cap_Shape[mush_subset$Cap_Shape == "c"] = "conical"
mush_subset$Cap_Shape[mush_subset$Cap_Shape == "x"] = "convex"
mush_subset$Cap_Shape[mush_subset$Cap_Shape == "f"] = "flat"
mush_subset$Cap_Shape[mush_subset$Cap_Shape == "k"] = "knobbed"
mush_subset$Cap_Shape[mush_subset$Cap_Shape == "s"] = "sunken"

mush_subset$Cap_Color <- as.character(mush_subset$Cap_Color)
mush_subset$Cap_Color[mush_subset$Cap_Color == "n"] <- "brown"
mush_subset$Cap_Color[mush_subset$Cap_Color == "b"] <- "buff"
mush_subset$Cap_Color[mush_subset$Cap_Color == "c"] <- "cinnamon"
mush_subset$Cap_Color[mush_subset$Cap_Color == "g"] <- "gray"
mush_subset$Cap_Color[mush_subset$Cap_Color == "r"] <- "green"
mush_subset$Cap_Color[mush_subset$Cap_Color == "p"] <- "pink"
mush_subset$Cap_Color[mush_subset$Cap_Color == "u"] <- "purple"
mush_subset$Cap_Color[mush_subset$Cap_Color == "e"] <- "red"
mush_subset$Cap_Color[mush_subset$Cap_Color == "w"] <- "white"
mush_subset$Cap_Color[mush_subset$Cap_Color == "y"] <- "yellow"

mush_subset$Odor <- as.character(mush_subset$Odor)
mush_subset$Odor[mush_subset$Odor == "a"] <- "almond"
mush_subset$Odor[mush_subset$Odor == "l"] <- "anise"
mush_subset$Odor[mush_subset$Odor == "c"] <- "creosote"
mush_subset$Odor[mush_subset$Odor == "y"] <- "fishy"
mush_subset$Odor[mush_subset$Odor == "f"] <- "foul"
mush_subset$Odor[mush_subset$Odor == "m"] <- "musty"
mush_subset$Odor[mush_subset$Odor == "n"] <- "none"
mush_subset$Odor[mush_subset$Odor == "p"] <- "pungent"
mush_subset$Odor[mush_subset$Odor == "s"] <- "spicy"


mush_subset$Gill_Color <- as.character(mush_subset$Gill_Color)
mush_subset$Gill_Color[mush_subset$Gill_Color == "k"] <- "black"
mush_subset$Gill_Color[mush_subset$Gill_Color == "n"] <- "brown"
mush_subset$Gill_Color[mush_subset$Gill_Color == "b"] <- "buff"
mush_subset$Gill_Color[mush_subset$Gill_Color == "h"] <- "chocolate"
mush_subset$Gill_Color[mush_subset$Gill_Color == "g"] <- "gray"
mush_subset$Gill_Color[mush_subset$Gill_Color == "r"] <- "green"
mush_subset$Gill_Color[mush_subset$Gill_Color == "o"] <- "orange"
mush_subset$Gill_Color[mush_subset$Gill_Color == "p"] <- "pink"
mush_subset$Gill_Color[mush_subset$Gill_Color == "u"] <- "purple"
mush_subset$Gill_Color[mush_subset$Gill_Color == "e"] <- "red"
mush_subset$Gill_Color[mush_subset$Gill_Color == "w"] <- "white"
mush_subset$Gill_Color[mush_subset$Gill_Color == "y"] <- "yellow"

mush_subset$Habitat <- as.character(mush_subset$Habitat)
mush_subset$Habitat[mush_subset$Habitat == "g"] <- "grasses"
mush_subset$Habitat[mush_subset$Habitat == "l"] <- "leaves"
mush_subset$Habitat[mush_subset$Habitat == "m"] <- "meadows"
mush_subset$Habitat[mush_subset$Habitat == "p"] <- "paths"
mush_subset$Habitat[mush_subset$Habitat == "u"] <- "urban"
mush_subset$Habitat[mush_subset$Habitat == "w"] <- "waste"
mush_subset$Habitat[mush_subset$Habitat == "d"] <- "woods"

head(mush_subset)
```




